Hunting the Parent of the Orphan Stream: Identifying Stream Members \\ from Low-Resolution Spectroscopy

### 2012-12-03 2012-oss-mrs/Response_to_referee_report.txt

 - Section 5,
   "help determine both the nature of the host system before it's discovered"

In both cases, 'host system' has been changed to 'progenitor'.

2) In the second sentence, the statement ``left behind in a stream of stars' is somewhat misleading, since tidal streams will extend both forward along the progenitor's orbit as well as behind. This should be amended to something like ``are strewn along the progenitor's orbit in leading and trailing streams of stars.'

Response:
Manuscript edited as requested. Following sentence edited to minimise repeated words. Text now reads as 'Stars which were once gravitationally bound to the satellite are distributed along the progenitor's orbit in leading and trailing streams of stars. The kinematics of the stream..'

@@

Response:
Corrected.

Section 4.2

9) In the first paragraph, it is not clear from Figure 2 why this particular separation line is appropriate or optimal. More explanation
and references should be provided. Also, ``allowing multiple inclusive cuts' is somewhat confusing. Should this be ``imposing simultaneous 
cuts on several observables'?

We have added additional text explaining Figure 2, and why the adopted separation line is appropriate for this work.

``allowing multiple inclusive cuts'' was confusing. The final text for this paragraph now reads as '..we are also employing selections on multiple observables (kinematics, metallicity, proper motions) in order to refine our Orphan Stream giant sample.'

@@

10) In the last paragraph, the selection criteria used by Casey et al. (2012a) should be briefly summarized here.

Response:
The text now reads: 'Given the target selection, our sample is likely to contain many more dwarfs than giants (e.g. see \citet{Casey_et-al_2012} where a similar colour selection was employed).' 
Section 4.3

@@

11) ``fainted' should be ``fainter'.

Response:
Corrected this typo.

12) In Table 1, the ra and dec components of proper motion should be listed separately. Strong cross-stream proper motion components could 
also be used to remove or reduce membership probabilities for some stars.

Response:
The proper motion components in Table 1 are now listed separately. It seems there was an error in the original calculation for the total proper motion of OSS-1. Instead of the proper motion being added in quadrature, the uncertainties were added in quadrature for both the measurement and it's uncertainty (yielding a $mu_{tot} = 6.1 \pm 6.1$). The proper motions are actually (4.1 +/- 4.5, -34.0 +/- 4.5) in RA and Dec respectively. As OSS-1 was already classified as a low-probability member based on inconsistent metallicities between Ca line strengths and isochrone fitting, this makes no change to the probability assignment.

@@

13) Given the small number and significant color and magnitude spread among the highly probable members about the best-fitting isochrone in 
Figure 4, a distance uncertainty of only 1 kpc seems unlikely. How was this uncertainty estimated?

Response:
The distance uncertainty is now determined from propagating the metallicity uncertainty (0.2 dex). A revised distance uncertainty of 2.0 kpc is now reflected throughout the text.

@@

16) The last sentence of the first paragraph should also refer to
Grillmair (2006) main sequence fitting distance measurement.

Response:
Reference rightly added.

17) Given the title of the paper, does the new metallicity measurement help in any way to identify a possible progenitor? At the very least, 
a statement should be included to the extent that the orbit and the metallicity measurements rule out any of the classical dwarfs, and that 
if a bound progenitor still exists, then it must be an as yet undiscovered ultrafaint dwarf.

Response:
Added text: 'Given the stream orbit derived by \citet{Newberg_et-al_2010}, they excluded all possible known halo objects except for the dissolved star cluster, Segue 1. \citet{Simon_et-al_2011} obtained spectroscopy for six members in Segue 1 and found an extremely wide metallicity dispersion: from $<-3.4$ to $-1.63$ dex. On the basis of the extremely low metallicity in the cluster and the wide chemical dispersion, they conclude that Segue 1 is a disrupted dwarf spheroidal galaxy. Although the data presented here indicates the Orphan Stream progenitor is a disrupted dwarf spheroidal galaxy, we cannot reliably associate Segue 1 as the parent without additional observational data.'

@@

18) The maximum likelihood estimates of the stream velocity given by the authors are either 85.3 /pm 4.4 km/s, or 82.1 /pm 1.4 km/s. A simple
weighted mean of the values in Table 1 gives 83.0 km/s. This is very significantly different from the 105 km/s (using the author's V_gsr system) 
measured by Newberg et al. for the same part of the stream using BHB stars. This disagreement needs to be addressed, as it seriously undermines 
any confidence in the purity of even the ``high probability' sample.

Response:
This disagreement was of concern to the authors before the initial manuscript was submitted. We believe this discrepancy is now resolved. We quoted the galactocentric velocity of the stream to be 110 km/s as per Newberg et al. 2011. This was somewhat misleading. Newberg et al. 2011 quotes the mean galactocentric velocity in their conclusions, where this number was originally taken from. However, in Table 2 of Newberg et al. 2011 they find a galactocentric velocity of 101.4 km/s at Lambda_Orphan = 18.4 degrees. This is the closest detection to our fields, at Lambda_Orphan ~= 23 degrees. The referee will recall that this is ~95 km/s on our galactocentric velocity scale. There is a sharp velocity gradient along the stream, decreasing to 38.4 km/s at Lambda_Orphan = 36 degrees. Therefore, at Lambda_Orphan ~= 23 degrees, a galactocentric velocity of 85-90 km/s (or 80-85 km/s on our scale) is quite reasonable. This 'back of the envelope' calculation ignores the intrinsic uncertainty in BHB velocities, or the uncertainties quoted by Newberg et al 2011 of 101.4 km/s +/- 2.9 +/- 9.8 km/s.

Since this implies our initial kinematic selection should actually be between 65 <= V_GSR <= 125 km/s (on our scale), this required a re-analysis of stars in the V_GSR range between 65-75 km/s. No candidates had velocities between 125-135 km/s, thus none were excluded. We identified two more possible candidates that also met our dwarf/giant constraint, and measured their metallicities. These are medium-probability candidates (OSS 15, 20) and have been added to Table 1. Additional explanatory text has been added to the manuscript explaining the velocity results we find in comparison with Newberg et al in Section 4.4
We are glad this error has been spotted and corrected.

@@

19) Moreover, though it is somewhat difficult to say given small number statistics, the metallicity distribution of the high probability stars 
looks more similar to the that of the off-stream, field star sample of Newberg et al. than it does to that of the stream sample.

20) Newberg et al. find velocity dispersions in various parts of the Orphan stream of around 10 km/s. Here the case is made for a much lower 
velocity dispersion, even colder than that of globular cluster streams like Pal 5 and GD-1. It may well be that Newberg et al.'s velocities are
generally less accurate than the present sample. However, one would still have to think about how a stream that is 8-10 times broader than these 
globular cluster streams should be kinematically just as cold or colder. Unfortunately, the disagreement with previous Orphan velocity measurements,
and the fact that removing the weak signal stars makes this disagreement even larger while making the stream colder, undermines my confidence considerably.

21) In thinking about this, I have made use of a velocity versus metallicity plot, with appropriate uncertainties. This makes it much easier to 
see why the authors are getting the results they see, and I would strongly recommend including such a plot in the final manuscript.

Response to (19), (20), (21):
Two additional plots have been added to the manuscript. The first is a plot showing the velocity and metallicity of our highly probable candidates. The second is a histogram showing the metallicity distribution function for our sample, with a comparison to the [Fe/H] of BHB stars in Newberg et al. 2011. The velocity discrepancy with Newberg et al. (2011) has been resolved in (18), and additional discussion text has been added addressing the metallicity differences in Section 4.4.

@@

22) The last sentence of the 3rd paragraph is somewhat mysterious. Is it the random motions that must be perpendicular to the line of sight, or 
the mean motions? The latter follows from the simple fact that the stars are in a stream, and the stream is nearly perpendicular to our line of 
sight. Unless I'm missing something important, this sentence should be deleted.

Response:
Removed the last sentence. It was a bit ambiguous. Sorry about that. AC

### 2012-12-03 2012-oss-mrs/ms.tex
Re-submitted version after GDC and SK comments

\documentclass[10pt,apjl]{emulateapj}
\usepackage{amssymb,graphics} % for resolution symbol

@@

\author{Andrew R. Casey\altaffilmark{1,2}, Stefan C. Keller\altaffilmark{1}, Gary Da Costa\altaffilmark{1}, Elizabeth Maunder\altaffilmark{1}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; \email{acasey@mso.anu.edu.au}}
\altaffiltext{2}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, 77 Massachusetts Avenue, Cambridge, MA 02139, USA}

\begin{abstract}
We present coordinates and properties of candidate K giant members in the Orphan Stream, which have been identified from low-resolution spectroscopic data taken with AAOmega on the Anglo-Australian Telescope. From modest S/N spectra and independent cuts in photometry, kinematics, gravity and metallicity we yield self-consistent, highly probable stream members. We find a revised stream distance of $22.5\,\pm\,1.0$\,kpc near the celestial equator, and our kinematic signature peaks at $V_{GSR} = 82.1\,\pm\,1.4$\,km s$^{-1}$. The observed velocity dispersion of our most probable members is consistent with arising from the velocity uncertainties alone. This indicates that at least along the line-of-sight, the Orphan Stream is kinematically quite cold. Our data indicates an overall stream metallicity of [Fe/H] $= -1.63\,\pm\,0.19$\,dex which is more metal-rich than previously found and unbiased by spectral type. Furthermore, the significant metallicity dispersion displayed by our probable members, $\sigma = 0.56$\,dex, suggests that the unidentified Orphan Stream parent is a dSph satellite. We highlight likely members for high-resolution spectroscopic follow-up.
\end{abstract}

@@

The Milky Way stellar halo has partly formed through the accretion of satellites which are disrupted by tidal forces as they fall into the Milky Way's potential. Stars which were once gravitationally bound to the satellite are strewn along the progenitor's orbit in leading and trailing streams of stars. The kinematics of the stream are sensitive to the shape of the galactic potential, allowing us to constrain the Milky Way potential, and reconstruct the formation history of the Galaxy. The level of accreted substructure in the Milky Way has only recently become apparent through multi-band photometric surveys like the Sloan Digital Sky Survey (SDSS). The more prominent of the detectable substructures, like Sagittarius, have been well-studied. One of the more prominent \--- yet less studied \--- substructures is that of the Orphan Stream.

The Orphan Stream was independently detected by both \citet{Grillmair_2006} and \citet{Belokurov_et-al_2006}, and is distinct from other substructures in the halo. The stream stretches over $60\,^\circ$ in the sky, has a low surface brightness and a narrow stream width of only $\sim$2\,$^\circ$ and as the name suggests, the parent object largely remains a mystery. The stream extends past the celestial equator \--- outside the SDSS footprint \--- but has not been detected in existing southern surveys \citep{Newberg_et-al_2010}. Whilst the parent system remains elusive, significant effort has been placed on associating the stream with known Milky Way satellites \citep{Zucker_et-al_2006, Fellhaur_et-al_2007,Jin_Lynden_Bell_2007,Sales_et-al_2008}. In contrast, there has been relatively limited observational work on the Orphan Stream itself other than the original discovery papers \citep{Grillmair_Dionatos_2006, Belokurov_et-al_2006, Belokurov_et-al_2007} and the work of \citet{Newberg_et-al_2010}. This is largely to be expected given the absence of multi-band photometry in the southern sky and the low total luminosity of the stream, making it difficult to reliably separate Orphan Stream members from halo stars. Further understanding on the extent of the stream awaits the SkyMapper and Pan-STARRS photometric surveys \citep{Keller_et-al_2007, Hodapp_et-al_2004}.

As \citet{Sales_et-al_2008} point out, there is a natural observational bias towards more massive and recent mergers like Sagittarius. Consequently, the fainter end of this substructure distribution has yet to be fully recovered, or thoroughly examined. Interestingly, there are indications that  some fainter substructures like the Orphan Stream and the Palomar 5 tidal tails \citep{Odenkirchen_et-al_2009} have orbits which seem to be best-fit by Milky Way models with nearly 60\% less mass \citep{Newberg_et-al_2010} than generally reported by \citet{Xue_et-al_2008} and \citet{Koposov_et-al_2010}. Such a discrepancy in the mass of the Milky Way is troublesome. More complete photometric and kinematic maps of these low total luminosity streams may provide the best test as to whether this mass discrepancy is real, or an artefact of incomplete observations. Whilst the full spatial extent of the Orphan Stream remains incomplete, we can examine the detailed chemistry of its members and investigate the stream history, as well as make predictions about the nature of the progenitor.

In this letter we present a detailed, self-consistent analysis to identify K giant members of the Orphan Stream. Using our selection method we have catalogued the locations of nine highly probable Orphan Stream candidates, all worthy of high-resolution spectroscopic follow up. In the following section we outline our photometric target selection. In \S\ref{sec:observations} we describe the low-resolution spectroscopic observations. The data analysis, including stream identification, is discussed in \S\ref{sec:analysis} and in \S\ref{sec:conclusions} the conclusions, predictions and future work are presented.

@@

We have targeted K giant members of the Orphan Stream in order to investigate their detailed chemistry. Because K-giants are difficult to unambiguously detect from photometry alone, low-resolution spectroscopy is required to estimate stellar parameters and determine radial velocities. The Orphan Stream has a low spatial over-density on the order of $\sim$10\%, which makes it difficult to separate stream members from halo stars. However, there is a well described distance gradient along the stream \citep{Belokurov_et-al_2007, Newberg_et-al_2010} which provides an indication on where we should focus our spectroscopic efforts.

The Orphan Stream is closest to us in two locations on the edge of the SDSS footprint: at the celestial equator \citep{Belokurov_et-al_2007}, and along outrigger SEGUE Stripe 1540 \citep{Newberg_et-al_2010}. These two locations are unequivocally the best place to recover bright stream members. We have targeted two fields centered on $(\alpha, \delta) =$ (10:48:15, 00:00:00) and (10:48:15, $-$02:30:00), and employed a combination of colour cuts with the SDSS DR 7 \citep{Abazajian_et-al_2009} data set in order to identify likely K giants:
\begin{eqnarray}
0.6 <& (g-i)_0 &< 1.7 \\
-15(g-i)_0 + 27 <& g_0 &< -3.75(g-i)_0 + 22.5 \\
15  <& i_0  &< 18 
\end{eqnarray}

Given our colour selection we expect to recover giants and contaminating dwarf contaminants. Although the 2MASS $JHK$ colours can help to separate dwarfs and giants, our target K-giants stars are too faint to be detected in the 2MASS catalogue.

@@

The data were reduced using the standard \textsc{2DFDR} reduction pipeline\footnote{http://www.aao.gov.au/2df/aaomega/aaomega\_2dfdr.html}. After flat-fielding, throughput calibration for each fibre was achieved using the intensity of skylines in each fibre. The median flux of dedicated sky fibres was taken for sky subtraction, and wavelength calibration was performed using arc lamp exposures between science frames. Three thirty minute science exposures were median-combined to assist with cosmic ray removal. The median S/N obtained in the red arm for our fields is modest at 35 per pixel. With the presence of strong Ca\,\textsc{II} triplet lines in the red arm we are able to ascertain reliable radial velocities and reasonable estimates on overall metallicity \citep[][and references therein]{Starkenburg_et-al_2010}. Our spectral region also includes gravity-sensitive magnesium lines: Mg\,\textsc{I} at 8807\,{\AA}, and the Mg\,\textsc{I}\,b 3$p$-4$s$ triplet lines at $\sim$5178\,{\AA}. As we demonstrate in the next section, these lines are sufficient to discriminate dwarfs from giants even with weak signal.

@@

\section{Analysis}
\label{sec:analysis}

We have employed a combination of separate criteria to identify likely Orphan Stream members: kinematics, a giant/dwarf indication from Mg\,\textsc{I} lines, and consistent metallicities from isochrone fitting and the Ca\,\textsc{II} triplet lines. Each criteria is discussed here separately.

\subsection{Kinematics}
Radial velocities were measured by cross-correlating our normalised spectra against a K-giant synthetic template with a temperature of 4500\,K, $\log{g}$ = 1 and $[\mbox{Fe/H}] = -1.5$ across the range $845 \leq \lambda \leq 870$\,nm. Heliocentric velocities were translated to the galactic rest frame by adopting the local standard of rest velocity as 220\,km s$^{-1}$ towards $(l, b) = (53\,^\circ, 25\,^\circ)$ \citep{Kerr_Lynden-Bell_1986, Mihalas_Binney_1981}\footnote{Where $V_{GSR} = V_{\textsc{helio}} + 220\sin{l}\cos{b} + 16.5\times[\sin{b}\sin{25} + \cos{b}\cos{25}\cos{(l - 53)}]$}. \citet{Newberg_et-al_2010} presented a comprehensive kinematic dataset on the Orphan Stream using BHB stars and employed a slightly different galactocentric transformation. This systematic difference equates to our galactocentric rest frame velocities being on average $\sim$5\,km s$^{-1}$ lower than those of \citet{Newberg_et-al_2010}.

@@

Figure \ref{fig:velocities} shows a histogram of our galactocentric velocities, compared to the predicted smooth line-of-sight velocity distribution for this region from the Besan\c{c}on model \citep{Robin_et-al_2003}. We have selected predicted stars from the Besan\c{c}on model using the same criteria as outlined in \S\ref{sec:target-selection} after using the \citet{Jordi_et-al_2006} colour transformations. It is clear that our target selection has yielded mostly nearby thick disk dwarf stars with $V_{GSR} \approx -120$\,km s$^{-1}$. The expected kinematic signature for the Orphan Stream, as \citet{Newberg_et-al_2010} reported, is labelled in Figure \ref{fig:velocities}. There is no obvious sharp kinematic peak representative of the Orphan Stream in our sample. From kinematics alone, our targets appears largely indistinguishable from a smooth halo distribution.

In this region \citet{Newberg_et-al_2010} found the Orphan Stream to have $V_{GSR} = 110$ km s$^{-1}$ from BHB stars, approximately $105$\,km s$^{-1}$ on our scale. To isolate potential Orphan Stream members we have nominated a relatively wide kinematic selection region between $75 \leq V_{GSR} \leq 135$ km s$^{-1}$, which yields twenty-five Orphan Stream candidates. The typical uncertainty in our velocities is $\pm{}5.0$\,km s$^{-1}$, although for our fainter targets this quickly increases above 10\,km s$^{-1}$.

@@

We have measured the equivalent width of the gravity-sensitive Mg\,\textsc{I} line at 8807 \AA{} to distinguish dwarfs from giants \citep{Battaglia_Starkenburg_2012}. At a given temperature (or $g - r$) and metallicity, giant stars present narrower Mg\,\textsc{I} absorption lines than their dwarf contaminants. Given our target selection, our sample is likely to contain many more dwarfs than giants. In some cases no Mg\,\textsc{I} 8807 \AA{} line was apparent, so an upper limit was estimated based on the S/N. In these cases the candidate was considered a ``non-dwarf" because we cannot exclusively rule out a metal-poor sub-giant with this criteria alone. For these purposes we are only looking for a simple indication as to whether a star is likely a dwarf or not. Figure \ref{fig:ew-mg} illustrates the trend with EW$_{\lambda8807}$ against SDSS de-reddened\footnote{All magnitudes presented in this letter are de-reddened using the \citet{Schlegel_Finkbeiner_Davis_1998} dust maps.} $g - r$, illustrating the dominant upper dwarf branch we wish to exclude. The adopted dwarf/giant separation line is shown. Admittedly this selection criteria is not particularly strict, but we are allowing multiple inclusive cuts and taking the intersection of these criteria to identify Orphan Stream giants.

@@

This analysis was also performed using the total equivalent width of the Mg\,\textsc{I}\,b triplet lines. Both analyses were entirely consistent with each other: essentially the same candidate selection was found using both techniques. However, given slightly poorer signal at the Mg\,\textsc{I}\,b triplet, we were forced to adopt many more upper limits than when using the 8807\,{\AA} line. Because we classify all upper limits as being ``non-dwarfs" (i.e. potential OSS giants), we deduced a slightly larger candidate sample for the Mg\,\textsc{I}\,b analysis, which was primarily populated by upper limits. In conclusion, we found the Mg\,\textsc{I} line at 8807\,{\AA} appeared to be a more consistent dwarf discriminant given our weak S/N \--- particularly for our fainter stars \--- and thus we have used the 8807\,{\AA} Mg\,\textsc{I} selection throughout the rest of our analysis.

Our dwarf/giant separation line in Figure \ref{fig:ew-mg} yields 425 potential giants. This fraction of giants ($\sim$44\%) is much larger than that found by \citeauthor{Casey_et-al_2012} \citeyear{Casey_et-al_2012}a who employed a similar colour selection. Hence, our gravity criteria is extremely inclusive and susceptible to dwarf contamination. Upon taking the intersection of our kinematic and gravity selections, we deduce eighteen stars that appear likely Orphan Stream K giants.

@@

% Note to the reader:
% In the acknowledgements, the final line including the text "the University of Washington"
% line does not appear if I don't have \phn here.
\phn

\begin{thebibliography}{26}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

@@

@@

\bibitem[{{Grillmair} \& {Dionatos}(2006)}]{Grillmair_Dionatos_2006}
{Grillmair}, C.~J., \& {Dionatos}, O. 2006, \apjl, 643, L17

@@

>>>>

\documentclass[10pt,apjl]{emulateapj}
\usepackage{amssymb,graphics} % for resolution symbol
\usepackage{multirow} % deluxetable wasn't doing it for me.

@@

\author{Andrew R. Casey\altaffilmark{1,2}, Gary Da Costa\altaffilmark{1}, Stefan C. Keller\altaffilmark{1}, Elizabeth Maunder\altaffilmark{1}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; \email{acasey@mso.anu.edu.au}}
\altaffiltext{2}{Massachusetts Institute of Technology and Kavli Institute for
Astrophysics and Space Research, 77 Massachusetts Avenue,
Cambridge, MA 02139, USA}

\begin{abstract}
We present candidate K-giant members in the Orphan Stream which have been identified from low-resolution data taken with the AAOmega spectrograph on the Anglo-Australian Telescope. From modest S/N spectra and independent cuts in photometry, kinematics, gravity and metallicity we yield self-consistent, highly probable stream members. We find a revised stream distance of $22.5\,\pm\,2.0$\,kpc near the celestial equator, and our kinematic signature peaks at $V_{GSR} = 82.1\,\pm\,1.4$\,km s$^{-1}$. The observed velocity dispersion of our most probable members is consistent with arising from the velocity uncertainties alone. This indicates that at least along this line-of-sight, the Orphan Stream is kinematically cold. Our data indicates an overall stream metallicity of [Fe/H] $= -1.63\,\pm\,0.19$\,dex which is more metal-rich than previously found and unbiased by spectral type. Furthermore, the significant metallicity dispersion displayed by our most probable members, $\sigma(\mbox{[Fe/H]}) = 0.56$\,dex, suggests that the unidentified Orphan Stream parent is a dSph satellite. We highlight likely members for high-resolution spectroscopic follow-up.
\end{abstract}

@@

The Milky Way stellar halo has partly formed through the accretion of satellites that are disrupted by tidal forces as they fall into the Galaxy's potential. Stars which were once gravitationally bound to the satellite are distributed along the progenitor's orbit in leading and trailing streams of stars. The velocities of stars in the stream are sensitive to the shape of the dark matter halo, allowing us to constrain the Milky Way potential and reconstruct the formation history of the Galaxy. The level of accreted substructure in the Milky Way has only recently become apparent through multi-band photometric surveys like the Sloan Digital Sky Survey (SDSS). The more prominent of the detectable substructures, like Sagittarius, have been well-studied. One of the more prominent \--- yet less studied \--- substructures is that of the Orphan Stream.

% could add a sentence to the paragraph above discussing the various substructures that have been detected

The Orphan Stream was independently detected by both \citet{Grillmair_2006} and \citet{Belokurov_et-al_2006}, and is distinct from other substructures in the halo. The stream stretches over $60\,^\circ$ in the sky, has a low surface brightness, and a narrow stream width of only $\sim$2\,$^\circ$.  As the name suggests, the parent object largely remains a mystery. The stream extends past the celestial equator \--- outside the SDSS footprint \--- but has not been detected in existing southern surveys \citep{Newberg_et-al_2010}. Whilst the parent system remains elusive, significant effort has been placed on associating the stream with known Milky Way satellites \citep{Zucker_et-al_2006, Fellhaur_et-al_2007,Jin_Lynden_Bell_2007,Sales_et-al_2008}. In contrast, there has been relatively limited observational work on the Orphan Stream itself other than the original discovery papers \citep{Grillmair_2006, Belokurov_et-al_2006, Belokurov_et-al_2007} and the work of \citet{Newberg_et-al_2010}. This is largely to be expected given the absence of deep multi-band photometry in the southern sky and the low total luminosity of the stream. This makes it difficult to reliably separate Orphan Stream members from halo stars. Understanding the full extent of the stream awaits the SkyMapper and Pan-STARRS photometric surveys \citep{Keller_et-al_2007, Hodapp_et-al_2004}.

As \citet{Sales_et-al_2008} point out, there is a natural observational bias towards more massive and recent mergers like Sagittarius. Consequently, the fainter end of this substructure distribution has yet to be fully recovered, or thoroughly examined. Interestingly, there are indications that  some fainter substructures like the Orphan Stream and the Palomar 5 tidal tails \citep{Odenkirchen_et-al_2009} have orbits which seem to be best-fit by Milky Way models with nearly 60\% less mass \citep{Newberg_et-al_2010} than generally reported by \citet{Xue_et-al_2008} and \citet{Koposov_et-al_2010}. Such a discrepancy in the mass of the Milky Way is troublesome. More complete photometric and kinematic maps of these low total luminosity streams may provide the best test as to whether this mass discrepancy is real, or an artefact of incomplete observations. Whilst the full spatial extent of the Orphan Stream remains unknown, we can examine the detailed chemistry of its members, investigate the stream history, and make predictions about the nature of the progenitor.

In this paper we present a detailed, self-consistent analysis to identify K-giant members of the Orphan Stream. Using our selection method we have catalogued the locations of nine highly probable Orphan Stream candidates, all worthy of high-resolution spectroscopic follow up. In the following section we outline our photometric target selection. In \S\ref{sec:observations} we describe the low-resolution spectroscopic observations. The data analysis, including stream identification, is discussed in \S\ref{sec:analysis} and in \S\ref{sec:conclusions} the conclusions, predictions and future work are presented.

@@

We have targeted K-giant members of the Orphan Stream in order to investigate their detailed chemistry. Because K-giants are difficult to unambiguously detect from photometry alone, low-resolution spectroscopy is required to estimate stellar parameters and determine radial velocities. The Orphan Stream has a low spatial over-density on the order of $\sim$10\%, which makes it difficult to separate stream members from halo stars. However, there is a well described distance gradient along the stream \citep{Belokurov_et-al_2007, Newberg_et-al_2010} which provides an indication on where we should focus our spectroscopic efforts.

The Orphan Stream is closest to us in two locations on the edge of the SDSS footprint: at the celestial equator \citep{Belokurov_et-al_2007}, and along outrigger SEGUE Stripe 1540 \citep{Newberg_et-al_2010}. These two locations are unequivocally the best place to recover bright stream members. We have targeted two fields centered on $(\alpha, \delta) =$ (10:48:15, 00:00:00) and (10:48:15, $-$02:30:00), and employed a combination of colour cuts with the SDSS DR 7 \citep{Abazajian_et-al_2009} data set in order to identify likely K-giants:
\begin{eqnarray}
0.6 <& (g-i)_0 &< 1.7 \\
-15(g-i)_0 + 27 <& g_0 &< -3.75(g-i)_0 + 22.5 \\
15  <& i_0  &< 18 
\end{eqnarray}

Given our colour selection we expect to recover giants and contaminating dwarfs. Although the 2MASS $JHK$ colours can help to separate dwarfs and giants, our target K-giants stars are too faint to be detected in the 2MASS catalogue.

@@

The data were reduced using the standard \textsc{2DFDR} reduction pipeline\footnote{http://www.aao.gov.au/2df/aaomega/aaomega\_2dfdr.html}. After flat-fielding, throughput calibration for each fibre was achieved using the intensity of skylines in each fibre. The median flux of dedicated sky fibres was used for sky subtraction, and wavelength calibration was performed using ThAr arc lamp exposures taken between science frames. Three thirty minute science exposures were median-combined to assist with cosmic ray removal. The median S/N obtained in the red arm for our fields is modest at 35 per pixel, although this deteriorates quickly for our fainter targets. With the presence of strong Ca\,\textsc{II} triplet lines in the red arm we are able to ascertain reliable radial velocities and reasonable estimates on overall metallicity \citep[][and references therein]{Starkenburg_et-al_2010}. Our spectral region also includes gravity-sensitive magnesium lines: Mg\,\textsc{I} at 8807\,{\AA}, and the Mg\,\textsc{I}\,b 3$p$-4$s$ triplet lines at $\sim$5178\,{\AA}. As we demonstrate in the next section, these lines are sufficient to discriminate dwarfs from giants even with weak signal.

@@

\section{Analysis \& Discussion}
\label{sec:analysis}

We have employed a combination of separate criteria to identify likely Orphan Stream members: kinematics, a giant/dwarf indication from Mg\,\textsc{I} lines, and selecting stars with consistent metallicities derived from both isochrone fitting and the strength of the Ca\,\textsc{II} triplet lines. Each criteria is discussed here separately.

\subsection{Kinematics}
Radial velocities were measured by cross-correlating our normalised spectra against a K-giant synthetic template with a temperature of 4500\,K, $\log{g}$ = 1.5 and $[\mbox{M/H}] = -1.5$ across the range $845 \leq \lambda \leq 870$\,nm. Heliocentric velocities were translated to the galactic rest frame by adopting the local standard of rest velocity as 220\,km s$^{-1}$ towards $(l, b) = (53\,^\circ, 25\,^\circ)$ \citep{Kerr_Lynden-Bell_1986, Mihalas_Binney_1981}\footnote{Where $V_{GSR} = V_{HELIO} + 220\sin{l}\cos{b} + 16.5\times[\sin{b}\sin{25^\circ} + \cos{b}\cos{25^\circ}\cos{(l - 53^\circ)}]$}.

Figure \ref{fig:velocities} shows a histogram of our galactocentric velocities, compared to the predicted smooth line-of-sight velocity distribution for this region from the Besan\c{c}on model \citep{Robin_et-al_2003}. We have selected particles from the Besan\c{c}on model using the same criteria outlined in \S\ref{sec:target-selection} after employing the \citet{Jordi_et-al_2006} colour transformations. It is clear that our target selection has yielded mostly nearby disk dwarf stars with $V_{GSR} \approx -120$\,km s$^{-1}$.

@@

In a nearby ($\Delta\Lambda_{Orphan} \sim 4\,^\circ$) region of the stream, \citet{Newberg_et-al_2010} detected the Orphan Stream with a $V_{GSR} = 101.4$ km s$^{-1}$ from BHB stars. Differences in accounting for the local standard of rest between this work and \citet{Newberg_et-al_2010} means that this corresponds to approximately 95\,km s$^{-1}$ on our $V_{GSR}$ scale.  This is discussed further in \S\ref{sec:newberg}. The expected Orphan Stream kinematic peak is labelled in Figure \ref{fig:velocities}. There is no obvious sharp kinematic peak representative of the Orphan Stream in our sample. From kinematics alone, our targets appears largely indistinguishable from a smooth halo distribution. To isolate potential Orphan Stream members we have nominated a relatively wide selection criteria between $65 \leq V_{GSR} \leq 125$ km s$^{-1}$ (shown in Figure \ref{fig:velocities}), which yields 28 Orphan Stream candidates. The typical uncertainty in our velocities is $\pm{}5.0$\,km s$^{-1}$.

@@

We have measured the equivalent width of the gravity-sensitive Mg\,\textsc{I} line at 8807 \AA{} to distinguish dwarfs from giants \citep{Battaglia_Starkenburg_2012}. At a given temperature (or $g - r$) and metallicity, giant stars present narrower Mg\,\textsc{I} absorption lines than their dwarf counterparts. Given the target selection, our sample is likely to contain many more dwarfs than giants (e.g. see \citet{Casey_et-al_2012} where a similar colour selection was employed). In some cases no Mg\,\textsc{I} 8807 \AA{} line was apparent, so an upper limit was estimated based on the S/N of the spectra. In these cases the candidate was considered a ``non-dwarf" because we cannot exclusively rule out a metal-poor sub-giant with this criteria alone. For these purposes we are only looking for a simple indication as to whether a star is likely a dwarf or not.

Figure \ref{fig:ew-mg} illustrates the trend with $EW_{\lambda8807}$ against SDSS de-reddened\footnote{All magnitudes presented in this letter are de-reddened using the \citet{Schlegel_Finkbeiner_Davis_1998} dust maps.} $g - r$, illustrating the dominant upper dwarf branch we wish to exclude. Giant stars populate the lower, sparser branch. A separation line has been adopted to distinguish dwarfs from giants, and is shown in Figure \ref{fig:ew-mg}. If we were to place this line higher, the total number of true giant stars may increase, but the dwarf contamination rate will rise dramatically. A compromise must be made between the rate of giant recoverability and the dwarf contamination. Our  dwarf/giant separation line lies just below the main dwarf population. On its own, this dwarf/giant separation line would typically result in far too many dwarf contaminants. However, we are  employing selections on multiple observables (kinematics, metallicity, proper motions) in order to refine our Orphan Stream giant sample.

@@

This dwarf/giant separation method was also employed using the total equivalent width of the Mg\,\textsc{I}\,b triplet lines. Both analyses were entirely consistent with each other: essentially the same candidate selection was found using both techniques. However, given slightly poorer signal at the Mg\,\textsc{I}\,b triplet, we were forced to adopt many more upper limits than when using the 8807\,{\AA} line. Because we classify all upper limits as being ``non-dwarfs" (i.e. potential giants), we deduced a slightly larger candidate sample for the Mg\,\textsc{I}\,b analysis, which was primarily populated by upper limits. In conclusion, we found the Mg\,\textsc{I} line at 8807\,{\AA} appeared to be a more consistent dwarf discriminant given our weak S/N \--- particularly for our fainter stars. Thus, we have used the 8807\,{\AA} Mg\,\textsc{I} selection throughout the rest of our analysis.

Our dwarf/giant separation line in Figure \ref{fig:ew-mg} yields 425 potential giants. Upon taking the intersection of our kinematic and gravity selections, we find 20 stars that appear to be likely Orphan Stream giants.

@@

\vspace{2em}

\begin{thebibliography}{29}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

@@

\bibitem[{{Frebel} \& {Norris}(2011){Frebel}, \& {Norris}}]{Frebel_Norris_2011}
{Frebel}, A., \& Norris, J. E. 2011, arXiv/1102.1748

@@

@@

\bibitem[{{Simon} {et~al.}(2011){Simon}, {Geha}, {Minor}, {Martinez},
	{Kirby}, {Bullock}, {Kaplinghat}, 
	{Strigari}, {Willman}, {Choi}, {Tollerud}, \&
	{Wolf}, J.}]{Simon_et-al_2011}
{Simon}, J.~D., {et~al.} 2011, \apj, 733, 46S
